---
title: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Visualizing Network Data

When the size of data grows larger, it becomes harder to visualize them. We can visualize it but, the graph may be hard to understand. 

#### Graph layouts

Let's try and see lattice network and web-logs networks (blogs). For blog network, we will use AIDS data collected by Gopal.

```{r load_pkg, message=FALSE}
# Load packages
library(igraph)
library(sand)

```

```{r graph_1, eval=FALSE}
## Lattice Network
g.1 <- graph.lattice(c(5, 5, 5))
```

```{r graph_2, eval=FALSE}
## blogs
data(aidsblog)
summary(aidsblog)

aidsblog <- upgrade_graph(aidsblog)

# Setting graph option
igraph.options(vertex.size = 3,
               vertex.label = NA,
               edge.arrow.size = 0.5)

# Plotting
par(mfrow=c(1, 2))

plor(g.1, layout = layout.circle)
title("5x5x5 Lattice")

plot(aidsblog, layout = layout.circle)
title("Blog Network")

```

So if you run the code above, you will find lattice network graph much more pleasing than the blog network graph as there are lesser edges crossing the center.

<br>

Let's explore more layouts.

```{r graph_3, eval=FALSE}
## Fruchterman and Reingold
par(mfrow=c(1, 2))

plot(g.1, layout = layout.fruchterman.reingold)
title("5x5x5 Lattice")

plot(aidsblog, layout = layout.fruchterman.reingold)
title("Blog Network")
```

```{r graph_4, eval=FALSE}
## Kamada and Kawai
par(mfrow=c(1, 2))

plot(g.l, layout=layout.kamada.kawai)
title("5x5x5 Lattice")

plot(aidsblog, layout=layout.kamada.kawai)
title("Blog Network")
```

```{r graph_5, eval=FALSE}
## Tree structure
g.tree = graph.formula(1-+2,1-+3,1-+4,2-+5,2-+6,2-+7,3-+8,3-+9,4-+10) par(mfrow=c(1, 3))

igraph.options(vertex.size = 30,
               edge.arrow.size = 0.5,
               vertex.label = NULL) 

plot(g.tree, layout = layout.circle)
plot(g.tree, layout = layout.reingold.tilford(g.tree, circular=T))
plot(g.tree, layout = layout.reingold.tilford)
```

```{r graph_6, eval=FALSE}
## Bipartite graph
g.bip <- upgrade_graph(g.bip)

plot(g.bip,
     layout = -layout.bipartite(g.bip)[,2:1],
     vertex.size = 30,
     vertex.shape =ifelse(V(g.bip)$type, "rectangle", "circle"),
     vertex.color = ifelse(V(g.bip)$type, "red", "cyan"))

```


### Decorating Graph Layouts

We have seen the importance of relative positions of vertices and edges in visualizing network graphs. However, we should also consider implementing additional information, such as size, shape and color of vertices and edges, when they are available.

**Karate Club Network**

- Nodes  represent members of karate club observed by Zachary  
- Connection between two nodes represent social interations between two corresponding members  
- Club is splitted into two different clubs due to a dispute between head teacher and an administrator  
- Attribute information for this network:  
    - identification of the head teacher and administrator
    - membership in one of the two factions underlying the eventual split  
    - relative frequency of interations between members  
    
```{r graph_7, eval=FALSE}
library(igraphdata)
data(karate)

# For this example, I've set seed to 42.
set.seed(42)


# Reproducible layout set.seed(42)
l <- layout.kamada.kawai(karate)


# Plot undecorated first.
par(mfrow=c(1,2))
plot(karate, layout = l, vertex.label = NA)


# Now decorate, starting with labels
V(karate)$label <- sub("Actor ", "", V(karate)$name)


# Two leaders get shapes different from club members
V(karate)$shape <- "circle"
V(karate)[c("Mr Hi", "John A")]$shape <- "rectangle"


# Differentiate two factions by color
V(karate)[Faction == 1]$color <- "red"
V(karate)[Faction == 2]$color <- "dodgerblue"


# Vertex area proportional to vertex strength 
# (i.e., total weight of incident edges).
(karate)$size <- 4 * sqrt(graph.strength(karate))
V(karate)$size2 <- V(karate)$size * .5


# Weight edges by number of common activities
E(karate)$width <- E(karate)$weight


# Color edges by within/between faction.
F1 <- V(karate)[Faction==1]
F2 <- V(karate)[Faction==2]

E(karate)[ F1 %--% F1 ]$color <- "pink"
E(karate)[ F2 %--% F2 ]$color <- "lightblue"
E(karate)[ F1 %--% F2 ]$color <- "yellow"


# Offset vertex labels for smaller points (default=0).
V(karate)$label.dist <- ifelse(V(karate)$size >= 10, 0, 0.75)


# Plot decorated graph, using same layout.
plot(karate, layout = l)

```

**Lazega Laywwer Network**

Let's look at the Lazega Laywer. As stated before, the lazega lawyer data were collected for the purpose of studying cooperation among social actors in an organization, through exchange of various types of resoruce among them. The data is observed from a law firm consisting over 70 lawyers in three offices located in different cities.

To visualize the graph easily, we will set offices to different colors. Also, we will set practices (litigation and corporate) to different shapes and the set vertex size (shape size) proportional to the years lawyers were with the firm.

```{r graph_8, eval=FALSE}
library(sand)
data(lazega)

lazega <- upgrade_graph(lazega)

# Office location indicated by color
colbar <- c("red", "dodgerblue", "goldenrod")
v.colors <- colbar[V(lazega)$Office]

# Type of practice indicated by vertex shape
v.shapes <- c("circle", "square")[V(lazega)$Practice]

# Vertex size proportional to years with firm
v.size <- 3.5 * sqrt(V(lazega)$Years)

set.seed(42)

l <- layout.fruchterman.reingold(lazega)

plot(lazega, 
     layout = l,
     vertex.color = v.colors,
     vertex.shape = v.shapes,
     vertex.size = v.size)


    ```


### Visualizing Large Network

With the methods shown above, there are some limitations at visualizing a graph when the number of vertices (data) increases as they tend to be clustered.

```{r graph_9, eval=FALSE}
library(sand)

summary(fblog)

set.seed(42)

party.names <- sort(unique(V(fblog)$PolParty))
party.names

l <- layout.kamada.kawai(fblog)

party.nums.f <- as.factor(V(fblog)$PolParty)

party.nums <- as.numeric(party.nums.f)

plot(fblog,
     layout = l,
     vertex.label = NA,
     vertex.color = party.nums,
     vertex.size = 3)
```

Running the code above, you may see that the graph is clustered and hard to find relationships between vertices. To solve this, let's look at the French political blog network using DrL Algorithm.

```{r graph_10, eval=FALSE}
l <- layout.drl(fblog)
plot(fblog,
     layout = l,
     vertex.size = 5,
     vertex.label = NA,
     vertex.color = party.nums)
```

We can see that clusters are formed in the graph. However, if we find characteristics in the graph explicitly, possibly through graph partitioning methods, it may be useful to replace cluster vertices into a single vertex.

```{r grpah_11, eval=FALSE}
fblog.c <- contract.vertices(fblog, party.nums)
E(fblog.c)$weight <- 1
fblog.c <- simplify(fblog.c)
party.size <- as.vector(table(V(fblog)$PolParty))

plot(fblog.c,
     vertex.size = 5 * sqrt(party.size),
     vertex.label = party.names,
     vertex.color = V(fblog.c),
     edge.width = sqrt(E(fblog.c)$weight),
     vertex.label.dist = 1.5,
     edge.arrow.size = 0)

k.nbhds <- graph.neighborhood(karate, order=1)
sapply(k.nbhds, vcount)

k.1 <- k.nbhds[[1]]
k.34 <- k.nbhds[[34]]

par(mfrow=c(1,2))
plot(k.1,
     vertex.label = NA,
     vertex.color = c("red", rep("lightblue", 16)))

plot(k.34,
     vertex.label = NA,
     vertex.color = c(rep("lightblue", 17), "red"))

```

